---
title: "Bellabeat Case Study"
author: "Swathy Ajaya Kumar"
date: "2022-09-30"
output: html_document
---

*This Case Study is the Capstone Project for the Google Data Analytics Course. The ideas and recommendations presented here in no way represent the ideas and belief of Google or Bellabeat. This project is completed for learning purposes only and may be inaccurate. The views expressed here are my own.*

***Assumption:*** *I am a junior data analyst working on the marketing analyst team at Bellabeat, a high-tech manufacturer of health-focused products for women. I have been asked to focus on one of Bellabeat’s products and analyze smart device data to gain insight into how consumers are using their smart devices. The insights will help guide marketing strategy for the company. The result of this analysis is assumed to be presented  to the Bellabeat executive team along with high-level recommendations for Bellabeat’s marketing strategy.*



### About Bellabeat

Founded in 2013 by Urška Sršen and Sando Mur, Bellabeat is high-tech company that manufactures beautifully designed health-focused smart products for women. Bellabeat products empower women with knowledge about their own health and habits by collecting data on activity, sleep, stress,and reproductive health. 

### Our Products 

Our wearables and accompanying products monitor biometric and lifestyle data to help women better understand how their bodies work and make healthier choices.

1. **Bellabeat app:** 
    Provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This app connects to our line of smart wellness products mentioned below.

2. **Leaf:**
    Wellness trackers that can be worn as a bracelet, necklace, or clip.

3. **Time:**
    Wellness SmartWatch is timepiece with smart technology to track user activity, sleep, and stress. 

4. **Spring:**
    Smart Water Bottle that tracks daily water intake using smart technology to ensure that our users are appropriately hydrated throughout the day. 

5. **Bellabeat membership:**
    We offer subscription-based membership program for users which gives users 24/7 access to fully personalized guidance on nutrition, activity, sleep, health and beauty, and mindfulness based on their lifestyle and goals.

### SWOT Analysis - Analysing Strengths, Weaknesses, Opportunities and Threats

#### Strengths
1. **Product Line-up:**
    Bellabeat's versatile product line-up includes wearable jewellery trackers, smart water bottle and smartwatch which can all be linked to the an app which can actively track activity,sleep, stress, menstrual cycle, and mindfulness habits.

2. **Beautiful Designs**
    Our products are beautifully and elegantly designed by women, for women. 
    
3. **Affordable Price Point**
    Our products are affordably priced compared to competitor products. Smart watches are priced at $149.00 USD + tax, and come with free 3 months subscription to personalized wellness coaching.
    
#### Opportunities
1. **Introduce wireless charging**
     We could modify our products to be compatible with wireless charging. We could sell wireless charging pods which are compatible with all our products.
    
2. **Customization**
    Our products are all elegantly designed. We could add more funky, sporty and trendy looks to accomodate the taste of more women. Not all women prefer minimalistic designs. 

#### Weaknesses
1. **Replaceable batteries**
    Bellabeat products currently use long-lasting, replaceable batteries. They last about 3-6 months. Disposable batteries contribute to e-waste.
    
2. **Limited options for customization**
    There aren't many options to choose from while selecting the style of products, especially the Leaf line-up. This might turn women away from buying our products

#### Threats
1. **Data Safety**
    On the wake of Roe v Wade being overturned, many women are worried if their reproductive data would be shared with the authorities. Therefore, many women are hesitant to track period data. Period data is essential for our personalized wellness program. We have enabled Private Key Encryption (AES-256) security feature for our mobile app to protect women's reproductive information. Bellabeat must find ways to protect women's rights in changing political landscape. 

2. **Competitor Threat**
    Bellabeat face strict competition from Fitbit, Samsung and Apple Watch. These products have economies of scale and have considerable market share. We must consistently push our Unique Selling Point (wellness products for women) to capture more of ther market share.
    
  
## Exploring Growth Opportunities

## Business Task
Urška Sršen, cofounder and Chief Creative Officer of Bellabeat is the *primary stakeholder* of this case study. Through this study, we will analyze smart device fitness data which will help unlock new growth opportunities for Bellabeat. 

## Preparing Data 
We will use  [FitBit Fitness Tracker Data](https://www.kaggle.com/datasets/arashnic/fitbit), which is a Kaggle data set that contains personal fitness tracker data from thirty Fitbit users. All users have consented to the submission of data . The data set includes output for physical activity, heart rate, and sleep monitoring, daily activity, steps, and heart rate. This can be used to explore users’ habits.

### Install Packages

We are using R to analyse this dataset. The following summarizes the packages that are installed to help us analyze user data. 

```{r}
install.packages("tidyverse")
install.packages("dplyr")
install.packages("skimr")
install.packages("here")
install.packages("janitor")
install.packages("tidyr")
install.packages("ggplot2")
install.packages("lubridate")

library("tidyverse")
library("dplyr")
library("skimr")
library("tidyr")
library("ggplot2")
library("lubridate")
library("here")
library("janitor")

```

## Uploading Datasets

```{r}
dailyActivity <- read_csv("dailyActivity_merged.csv")
dailysleep <- read_csv("sleepDay_merged.csv")
dailyweight <- read_csv("weightLogInfo_merged.csv")

```

#### Cleaning up data


##### Removing empty rows and columns
```{r}
 remove_empty(dailyActivity, which = c("rows", "cols"))
 remove_empty(dailysleep, which = c("rows", "cols"))
 remove_empty(dailyweight, which = c("rows", "cols"))
```
##### Check number of users

```{r}
n_distinct(dailyActivity$Id)
n_distinct(dailysleep$Id)
n_distinct(dailyweight$Id)
```

There are 33, 24 and 8 distinct user IDs reported for daily activity, sleep and weight log respectively. 

##### Check for duplicates

```{r}
sum(duplicated(dailyActivity))
sum(duplicated(dailysleep))
sum(duplicated(dailyweight))

```

The Sleep log has 3 duplicate values. We would have to remove them.

```{r}
dailysleep <- dailysleep %>% 
  distinct() %>% 
  drop_na()
```

```{r}
sum(duplicated(dailysleep))
```
##### Standard column names
Using the clean names function to standardize column names. 

```{r}

dailyActivity <- clean_names(dailyActivity)
dailysleep <- clean_names(dailysleep)
dailyweight <- clean_names(dailyweight)

head(dailyActivity)
head(dailysleep)
head(dailyweight)
```

##### Check the structure of each table 


```{r}
str(dailyActivity)
str(dailysleep)
str(dailyweight)
```

The datatype of the date column is set to "character". Changing the datatype to "Date"

```{r}

dailyActivity$activity_date <- as.Date(dailyActivity$activity_date, "%m/%d/%y")
dailysleep$sleep_day <- as.Date(dailysleep$sleep_day, "%m/%d/%y")
dailyweight$date <- as.Date(dailyweight$date, "%m/%d/%y")

head(dailyActivity)
head(dailyweight)
head(dailyweight)
```
### Cleaning conclusion
We removed empty rows and columns, checked distinct number of users for each table, checked for and removed duplicate values, standardized column names and changed the data type of the date columns of all tables from "character (chr)" to "date". 
This makes it easier for us to work with the data. 

## Summary for each data frame

Gathering some quick, basic statistical insights to each data frame helps us frame key questions we want answered in this case study. 

```{r}
summary(dailyActivity)
```

```{r}
summary(dailyweight)
```
```{r}
summary(dailysleep)
```


## Key Questions -  Analysing Data

Using available data, we will try to answer the following questions: 

1. On what days of the week were users more active? On what days were they sedentary?

2. How much time do users spend in bed compared to actual minutes asleep?

3. What is the relationship between users' calories burnt and sleep pattern?


### 1. On what days of the week were users more active? On what days were they sedentary?

We have the daily activity data and sleep data of 33 users available to us. First, we have to create a new column titled
```{r}
activity_analysis<- dailyActivity %>% 
  mutate(total_active_minutes= very_active_minutes+fairly_active_minutes+lightly_active_minutes)
head(activity_analysis)
```


```{r}
activity_analysis$weekday = weekdays(activity_analysis$activity_date)
head(activity_analysis)
```

```{r}
active_minutes_data <- activity_analysis %>%
  group_by(weekday) %>%
  summarize(total_active_minutes = mean(total_active_minutes)) %>%
  arrange(desc(total_active_minutes))

head(active_minutes_data)

```

```{r}
ggplot(active_minutes_data, aes(x=weekday, y=total_active_minutes)) +
  geom_bar(stat="identity", aes(x=weekday, y=total_active_minutes, fill=total_active_minutes)) + 
  scale_fill_gradient(low="#BF3A30", high="#864BA2", guide="none") +
  xlab("") +
  ylab("Total Active Minutes") +
  labs(title="Users' Total Weekly Active Minutes") +
  theme(aspect.ratio = 1,
        text=element_text(family="Ariel", size=15),
        plot.title=element_text(hjust=0.05, face="bold", size=18, color="#7E413A", margin=margin(b=19)),
        axis.title.x=element_text(margin=margin(t=5)),
        axis.title.y=element_text(size=9, color="#BA6C65", margin=margin(r=)),
        axis.text.x=element_text(size=8, color="#5C5C5C", margin=margin(t=-2)),
        axis.text.y.left=element_text(color="#BA6C65"),
        axis.ticks.length.x=unit(0, "cm"),
        axis.ticks.length.y=unit(0, "cm"),
        panel.grid.major.x=element_blank(),
        panel.grid.major.y=element_line(color="#FCB59C"),
        legend.title=element_text(color="#333333"),
        legend.text=element_text(color="#333333"),
        panel.background=element_rect(fill="white", color="white")) +
scale_x_discrete(guide = guide_axis(n.dodge=2))

```



Similarly, let's find out which days were users least active:

```{r}
activity_analysis %>%
  group_by(weekday) %>%
  summarize(sedentary_minutes = mean(sedentary_minutes))  %>%
  arrange(desc(sedentary_minutes))
```

Users were *most active on Thursdays* and *least active on Saturdays.*


## 2. How much time do users spend in bed compared to actual minutes asleep?

National Sleep Foundation guidelines advise that healthy adults need between 7 and 9 hours of sleep per night. It's important to wind down and get the most out of your time in bed. Next, we will compare the time spend in bed to actual minutes asleep.


```{r}
ggplot(data = dailysleep) +
  geom_point(mapping = aes(x = total_minutes_asleep, y = total_time_in_bed), color =  'magenta') +
  geom_smooth(mapping = aes(x = total_minutes_asleep, y = total_time_in_bed), color = 'black') +
  labs(title = "Total Minutes Asleep vs Total Time in Bed")
```

## 3. What is the relationship between users' calories burnt and sleep pattern?

```{r}
names(dailysleep)[names(dailysleep) == 'sleep_day'] <- 'date'
names(dailyActivity)[names(dailyActivity) == 'activity_date'] <- 'date'
head(dailysleep)
head(dailyActivity)

```

We will perform a full INNER JOIN to combine sleep data and activitie data. By doing this, we ensure that only the most relavent data entries are considered. 

```{r}
dailycal_dailysleep<- merge(dailyActivity, dailysleep, by=c ('id', 'date'), all = FALSE)
head(dailycal_dailysleep)
```

We can also create a new data frame from the existing one with just the sleep and calories columns.

```{r}
dailycal_dailysleep_merged<- dailycal_dailysleep[, c("id", "date", "calories","total_minutes_asleep")]
head(dailycal_dailysleep_merged)

```
We'll now perform a correlation test to understand if total minutes asleep and total calories burnt have positive correlation or negative correlation, and to what degree they're correlated. 

```{r}
cor.test(dailycal_dailysleep_merged$total_minutes_asleep, dailycal_dailysleep_merged$total_minutes_asleep, method=c("pearson"))

```
The correlation test says there significant correlation between total minutes asleep and total calories burnt. To help visualise this correlation, let's look at the scatterplot.

```{r}
ggplot(data = dailycal_dailysleep_merged) +
  geom_point(mapping = aes(x = total_minutes_asleep, y = calories), color =  'magenta') +
  geom_smooth(mapping = aes(x = total_minutes_asleep, y = calories), color = 'black') +
  labs(title = "Total Minutes Asleep vs Total Calories burnt")
```

# Analysis - Summary

1. Users were more active on Thursdays and were least active on Saturdays.
2. Users fell asleep as soon as they hit the bed. There wasn't much difference between time in bed vs actual time asleep. 
3. Users who sleep more tend to burn more calories 

# Recommendations:

First, we should acknowledge that the data set used for this case study is small. Nevertheless, we gained significant insights from this study. Following are the recommendations for Bellabeat:

### 1.    Notifications:
Sometimes a nudge in the right direction is all we need to stay on track of our fitness goals. We saw from user data that users tend to be lethargic on Saturdays. The Bellabeat fitness app can send out push notifications to motivate users by reminding them of how well they did on weekdays. Constant reminders like this will ensure user traffic

### 2.    Good night's sleep
The devices can track user's sleep data and send vibration alerts if they're spending too much time in bed without sleeping.  Sleep data collected could include resting heart rate while sleeping, restlessness etc and provide insights (articles, videos, sleep podcasts, etc) the next day for a good night's sleep. 
### 3.    Rewards for tracking weight and other data
Too few users tracked their weight. Bellabeat can encourage users to keep track of their data, especially those which require manual input, like weight. It can be badges, listing personal best weeks, coupons, access to premium content or even discounts on Bellabeat subscription.  





